\section*{Acknowledgments}

Before presenting the work carried out during the months of this thesis, I would like to take a moment to thank the people without whom this endeavour would never have come to fruition.

First and foremost, I would like to express my gratitude to my supervisor, Luca Roversi. 
Throughout my academic journey, he has undoubtedly been the most influential figure, reigniting a passion that had begun to wane. 
His guidance in preparing this thesis allowed me to delve into topics of personal interest, granting me considerable autonomy while never leaving me feeling ``abandoned''. 
Thank you, not only for being an exceptional mentor but also for the profound humanity with which you support your students throughout their journey.

Despite my efforts, many of the results achieved would not have been possible with such ease if not for the unwavering support of my parents. 
Thanks to them, I could focus solely on my studies, taking the necessary time and making my mistakes. 
Not everyone has this opportunity, and I am deeply grateful for their constant presence. 
I cannot overlook my brother either—thank you for always bringing a smile to my face and for never knowing the answers to the math questions I ask you; you remind me that it's perfectly human not to have all the answers right away.

A special thanks to Silvia, who has patiently supported and stood by me since the beginning of university. 
You are perhaps the person who has most witnessed my frustration when I struggled to reach my goals. 
You have been, and continue to be my calm haven, and I hope you always will be.

I would also like to extend my gratitude to the Montano and Rozzino families, who represent a ``second family'' I can always rely on, no matter what.

Thanks to all the friends who accompanied me along the way, both long-time and recent, each contributing uniquely to this important milestone in my life.

Lastly, but by no means least, I want to thank my grandparents—this thesis is dedicated to you.